\begin{abstract}
Scientific knowledge graphs (KGs) encode entities and typed relations across
domains such as physics, astronomy, and materials science, yet they remain
incomplete: missing edges and entities limit downstream reasoning.  We introduce
\emph{Harmony}, a framework that treats theory discovery as the search for KG
mutations---new edges or entities---that maximise a composite quality metric.
The \emph{Harmony score} combines four complementary signals: \textbf{compressibility}
(minimum description length proxy), \textbf{coherence} (path-semantic consistency),
\textbf{symmetry} (intra-type behavioural consistency),
and \textbf{generativity} (link-prediction learnability via DistMult), with an
optional \textbf{frequency} component for hybrid scoring.  An LLM
proposer generates candidate theory-level propositions, which are validated,
scored, and archived in a MAP-Elites quality-diversity grid across an
island-model search topology.  We evaluate on seven KG domains (five
hand-curated, two Wikidata-sourced) with 10-seed cross-validation and four
external KGE baselines (DistMult, TransE, RotatE, ComplEx).  Factor
decomposition isolates the contributions of the LLM proposer, Harmony scoring,
and quality-diversity archive.  A regime characterisation analysis reveals where
Harmony's design aligns with KG structure: on Wikidata Materials,
Harmony-guided proposals show the largest directional improvement over
DistMult-alone on both Hits@10 ($0.34$ vs.\ $0.32$) and MRR ($0.12$
vs.\ $0.11$), though differences do not reach statistical significance
at $\alpha = 0.05$ and the effect size $|\delta_C|\,\leq\,0.12$ falls in
the ``negligible'' range per \citet{cliff1993dominance}.  A
pre-registered calibration gate passed on the two calibration domains
($31$--$65\%$ over frequency) but did not transfer to the discovery
domains, a regime-finding rather than a refutation.
Backtesting against held-out edges shows that proposals do not reproduce
known edges, serving as a negative control consistent with theory-level
rather than interpolation behaviour; a typed-neighbourhood soft-match
diagnostic suggests proposals tend toward structurally adjacent knowledge.
We frame Harmony as a methodology for theory-shaped proposal generation;
assessing whether produced propositions constitute novel scientific
discoveries requires human expert evaluation, which we defer to future work.
\end{abstract}
