\section{Dataset Statistics and Reproducibility}\label{app:datasets}

Table~\ref{tab:reproducibility} provides a unified summary of all seven KG
domains, including data sources and construction methods.  This addresses the
need for consistent dataset definitions across hand-curated and Wikidata-sourced
KGs.

\begin{table*}[h]
\centering
\caption{Unified reproducibility table for all seven KG domains.  All KGs use
the shared seven-relation type vocabulary.  Entity and edge counts are computed
programmatically from the KG builders.}\label{tab:reproducibility}
\begin{tabular}{lccccl}
\toprule
Domain & Entities & Edges & Entity types & Edge types & Source \\
\midrule
Linear algebra     & 50  & 75  & 2 & 7 & Hand-built \\
Periodic table     & 153 & 326 & 4 & 2 & Hand-built \\
Astronomy          & 41  & 39  & 6 & 6 & Hand-built \\
Physics            & 43  & 46  & 6 & 5 & Hand-built \\
Materials science  & 49  & 53  & 6 & 5 & Hand-built \\
Wikidata Physics   & 252 & 814 & 12 & 7 & Wikidata \\
Wikidata Materials & 253 & 806 & 18 & 7 & Wikidata \\
\bottomrule
\end{tabular}
\end{table*}

\noindent All statistics are computed programmatically from the KG builders
via \texttt{analysis/reproducibility\_table.py}.  The dataset splitting protocol
(Section~\ref{sec:experiments}) applies identically to all seven domains:
10\% hidden, then 80/10/10 train/val/test on the remaining 90\%.

\section{Ablation Details}\label{app:ablation}

The ablation study (Table~\ref{tab:ablation}) uses the linear algebra KG with
$n_{\text{bootstrap}} = 200$ samples.  For each ablation variant, one weight is
set to zero and the remaining three are renormalised to sum to 1.  Bootstrap 95\%
confidence intervals are computed via the percentile method on the mean Harmony
score.

\paragraph{Weight sensitivity.}
We evaluate six weight configurations from the calibration gate grid
($\alpha \in \{0.3, 0.5, 0.7\}$, $\beta \in \{0.1, 0.3\}$,
$\gamma = \delta = 0.25$).  All configurations show Harmony~$>$~frequency
baseline, with $\alpha = 0.5, \beta = 0.3$ yielding the highest mean Harmony
score.  This suggests that a moderate compressibility weight combined with
non-trivial coherence weight best captures the structure of our curated KGs.

\section{Proposal Validation Rules}\label{app:validation}

The deterministic validator enforces three rules:
\begin{enumerate}[nosep]
  \item \textbf{Text length}: \texttt{claim}, \texttt{justification},
    and \texttt{falsification\_condition} must each be $\geq 10$ characters.
    \texttt{kg\_domain} must be $\geq 3$ characters (controlled vocabulary,
    not free text).
  \item \textbf{Type-specific fields}: \texttt{ADD\_EDGE} requires
    \texttt{source\_entity}, \texttt{target\_entity}, and \texttt{edge\_type};
    \texttt{ADD\_ENTITY} requires \texttt{entity\_id} and \texttt{entity\_type};
    \texttt{REMOVE\_EDGE} requires \texttt{source\_entity},
    \texttt{target\_entity}, and \texttt{edge\_type};
    \texttt{REMOVE\_ENTITY} requires \texttt{entity\_id}.
  \item \textbf{Edge type validity}: \texttt{edge\_type} must be one of the seven
    valid \texttt{EdgeType} names.
\end{enumerate}

\section{Full Proposal Examples}\label{app:proposals}

Below are three complete proposal records from the astronomy archive, showing
all fields including justification and falsification conditions.

\paragraph{Proposal 1: Stellar nucleosynthesis $\to$ heavy element abundance.}
\begin{itemize}[nosep]
  \item \textbf{Type}: \texttt{ADD\_EDGE}
  \item \textbf{Source}: \texttt{stellar\_nucleosynthesis}
  \item \textbf{Target}: \texttt{heavy\_element\_abundance}
  \item \textbf{Edge type}: \texttt{explains}
  \item \textbf{Claim}: ``Stellar nucleosynthesis explains the observed abundance
    pattern of heavy elements in planetary nebulae.''
  \item \textbf{Justification}: ``The s-process and r-process nucleosynthesis
    pathways in AGB stars and supernovae produce characteristic abundance
    patterns that match spectroscopic observations of planetary nebulae.''
  \item \textbf{Falsification}: ``Discovery of heavy element abundance patterns
    in planetary nebulae inconsistent with any known nucleosynthesis pathway
    would falsify this claim.''
\end{itemize}

\paragraph{Proposal 2: Mass--luminosity relation derivation.}
\begin{itemize}[nosep]
  \item \textbf{Type}: \texttt{ADD\_EDGE}
  \item \textbf{Source}: \texttt{hydrostatic\_equilibrium}
  \item \textbf{Target}: \texttt{mass\_luminosity\_relation}
  \item \textbf{Edge type}: \texttt{derives}
  \item \textbf{Claim}: ``The mass--luminosity relation derives from hydrostatic
    equilibrium in main sequence stars.''
  \item \textbf{Justification}: ``Balancing gravitational pressure against
    radiation pressure in the stellar core, combined with opacity-dependent
    energy transport, yields $L \propto M^{3.5}$ for main sequence stars.''
  \item \textbf{Falsification}: ``A main sequence star population where
    luminosity is uncorrelated with mass would disprove this derivation.''
\end{itemize}

\paragraph{Proposal 3: Magnetar as new entity.}
\begin{itemize}[nosep]
  \item \textbf{Type}: \texttt{ADD\_ENTITY}
  \item \textbf{Entity ID}: \texttt{magnetar}
  \item \textbf{Entity type}: \texttt{celestial\_object}
  \item \textbf{Claim}: ``Magnetar generalises the neutron star category with
    extreme magnetic field properties ($B > 10^{14}$~G).''
  \item \textbf{Justification}: ``Magnetars are observationally distinct from
    ordinary neutron stars due to their ultra-strong magnetic fields, which power
    soft gamma repeaters and anomalous X-ray pulsars.''
  \item \textbf{Falsification}: ``Evidence that magnetar-attributed emissions
    originate from non-magnetic mechanisms would undermine this classification.''
\end{itemize}

\section{LLM Prompt Templates}\label{app:prompts}

We include the exact prompt templates used for proposal generation.  Both modes
share a common preamble with KG statistics, strategy instruction, top proposals,
and recent failures.

\paragraph{Free mode (default).}
The free-mode prompt shows a sample of up to 20 entity IDs from the KG to
ground the LLM without over-constraining it:
\begin{quote}\small\ttfamily
You are a theory-discovery agent for knowledge graph research.\\
Knowledge Graph: domain='\{domain\}', entities=\{N\}, edges=\{M\}\\[4pt]
Strategy: \{REFINEMENT|COMBINATION|NOVEL\} --- \{strategy description\}\\[4pt]
Top proposals so far: \{top 3 proposals or "None yet"\}\\
Recent validation failures: \{up to 5 failure messages or "None"\}\\[4pt]
EXAMPLE ENTITY IDs from this KG (showing K of N): \{entity\_1\}, \{entity\_2\}, ...\\
VALID EDGE TYPES: depends\_on, derives, equivalent\_to, maps\_to, explains,
contradicts, generalizes\\
IMPORTANT: source\_entity and target\_entity MUST be exact entity IDs from this
KG.\\[4pt]
Return ONLY a JSON object (no extra text) with fields: id, proposal\_type,
claim, justification, falsification\_condition, kg\_domain, source\_entity,
target\_entity, edge\_type, entity\_id, entity\_type
\end{quote}

\paragraph{Constrained mode (stagnation recovery).}
When an island stagnates ($S = 5$ generations without valid proposals), the
prompt switches to constrained mode, which enumerates \emph{all} valid entity
IDs and edge type names explicitly:
\begin{quote}\small\ttfamily
... [same preamble] ...\\[4pt]
VALID ENTITY IDs (use EXACTLY as written): \{all entity IDs\}\\
VALID EDGE TYPES (use EXACTLY as written): depends\_on, derives, equivalent\_to,
maps\_to, explains, contradicts, generalizes
\end{quote}

\section{Proposal Failure Rate Statistics}\label{app:failure-rate}

Figure~\ref{fig:convergence} shows the valid proposal rate converging to
$\geq 0.50$ by generation~10 across all discovery domains.  The initial failure
rate (generations 1--3) is typically 60--80\%, dominated by entity grounding
errors (referencing entities not in the KG).  The entity sample in free-mode
prompts (up to 20 entities) and the stagnation recovery mechanism
(Section~\ref{sec:search}) together reduce failures to $<$30\% by generation~10.
Constrained-mode prompts achieve $\geq$95\% validity but produce less diverse
proposals.

\section{Code and Data Availability}\label{app:code-data}

Source code and all experimental artifacts are publicly available:
\begin{itemize}[nosep]
  \item \textbf{Code repository}: anonymised for review; will be released upon
    acceptance.
  \item \textbf{Data archive}: Zenodo anonymized draft record
    (\texttt{https://zenodo.org/deposit/18795697}),
    containing all KG datasets, checkpoints, and generated proposals.
\end{itemize}

\section{Statistical Methods}\label{sec:stat-methods}

All pairwise comparisons between methods use the following statistical tests,
computed over 10 random seeds per domain:

\paragraph{Paired bootstrap CI.}
Given paired score vectors $\mathbf{a} = (a_1, \ldots, a_{10})$ and
$\mathbf{b} = (b_1, \ldots, b_{10})$, we resample $B = 2000$ bootstrap
differences $d^{(b)} = \frac{1}{10}\sum_i (a_{\pi_i} - b_{\pi_i})$ (sampling
with replacement) and report the 95\% percentile CI $[d_{0.025}, d_{0.975}]$.
The two-sided $p$-value is $p = 2 \cdot \min\!\bigl(\hat{P}(d^{(b)} \leq 0),\;
\hat{P}(d^{(b)} \geq 0)\bigr)$.

\paragraph{Cliff's delta.}
A non-parametric effect size: $\delta_C = \frac{1}{n^2}\sum_{i,j}
\text{sign}(a_i - b_j) \in [-1, 1]$.  We interpret $|\delta_C| < 0.147$ as
negligible, $< 0.33$ as small, $< 0.474$ as medium, and $\geq 0.474$ as large
\citep{cliff1993dominance}.

\paragraph{Permutation test.}
We randomly swap labels between methods $10{,}000$ times to compute a Monte
Carlo estimate of the $p$-value under the null hypothesis of no difference.

\section{Frequency Baseline Analysis}\label{app:freq}

For a KG with edge-type frequency vector $\mathbf{p} = (p_1, \ldots, p_7)$, the
frequency heuristic predicts the most common edge type for every unobserved
entity pair.  The expected Hits@$K$ of this strategy satisfies:
\[
  \text{Hits@}K \geq p_{\max} = \max_i p_i,
\]
since for any test edge with the dominant type, the frequency prediction ranks
the correct type first.  In our domains, $p_{\max}$ ranges from 0.22 (Wikidata
Materials, highest entropy) to 0.38 (Physics, lowest entropy), providing a lower
bound on frequency Hits@10.

The Shannon entropy $H(\mathbf{p}) = -\sum_i p_i \log_2 p_i$ quantifies
edge-type diversity; lower entropy indicates more peaked distributions where
frequency is more informative.  A regression of frequency Hits@10 on
$H(\mathbf{p})$ across our seven domains shows a negative correlation,
confirming the information-theoretic explanation.

\section{Wikidata Ablation}\label{app:wikidata-ablation}

Per-component ablation on the Wikidata Physics and Wikidata Materials domains
uses $n_{\text{bootstrap}} = 200$ resamples.  On the denser Wikidata KGs,
coherence shows a larger contribution than on the sparser hand-curated domains,
consistent with the higher triangle density in larger graphs.  Compressibility
and symmetry contributions are broadly consistent across all domains.
Generativity remains the single largest contributor, though the Harmony-3
analysis (Section~\ref{sec:results}) shows that the structural components alone
provide useful signal.

\section{Hyperparameter Settings}\label{app:hyperparams}

Table~\ref{tab:hyperparams} lists all hyperparameters used in the experiments.

\begin{table}[h]
\centering
\caption{Hyperparameter settings.}\label{tab:hyperparams}
\begin{tabular}{llc}
\toprule
Component & Parameter & Value \\
\midrule
\multirow{5}{*}{Harmony metric}
  & $\alpha$ (compressibility) & 0.25 \\
  & $\beta$ (coherence) & 0.25 \\
  & $\gamma$ (symmetry) & 0.25 \\
  & $\delta$ (generativity) & 0.25 \\
  & $\varepsilon$ (frequency) & 0.0 \\
\midrule
\multirow{6}{*}{DistMult}
  & Embedding dimension & 50 \\
  & Training epochs & 100 \\
  & Margin & 1.0 \\
  & Learning rate & 0.01 \\
  & Negative samples & 5 \\
  & Mask ratio & 0.20 \\
\midrule
\multirow{5}{*}{Search loop}
  & Islands & 4 \\
  & Population per island & 5 \\
  & Generations & 20 \\
  & Migration interval & 10 \\
  & Temperatures & $\{0.3, 0.3, 0.8, 1.2\}$ \\
\midrule
\multirow{2}{*}{Stagnation}
  & Trigger generations ($S$) & 5 \\
  & Recovery generations ($R$) & 3 \\
\midrule
\multirow{2}{*}{MAP-Elites}
  & Grid size & $5 \times 5$ \\
  & Descriptors & simplicity, Harmony gain \\
\midrule
Value function
  & $\lambda$ (cost penalty) & 0.1 \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{Compute resources.}
All experiments were run on a single Apple M2 Pro (10-core CPU, 32\,GB RAM, no
GPU).  Main experiments use \texttt{gpt-oss:20b} via Ollama ($\sim$10~min per
domain, 20~generations).  Factor decomposition baselines (LLM-only, No-QD) use
\texttt{mlx-community/Qwen3.5-35B-A3B-4bit} via \texttt{mlx\_lm} at $\sim$46~tok/s,
completing all three configurations across five domains in $\sim$2.5~hours.
The total compute for the seven reported domains is under 2~CPU-hours for the
main experiments; factor decomposition adds $\sim$2.5~CPU-hours.
KGE baseline training (TransE, RotatE, ComplEx across 5 domains $\times$ 10
seeds) adds approximately 30~minutes.  Statistical tests (bootstrap CI,
permutation tests) add $<$10~minutes.
