\section{Discussion}\label{sec:discussion}

\paragraph{Compressibility--generativity tension.}
Adding edges to a KG typically \emph{reduces} compressibility (the BFS spanning
fraction drops as cross-edges are introduced) while potentially \emph{improving}
generativity (more training signal for DistMult).  This tension is by design:
the Harmony metric rewards proposals that improve link-prediction learnability
without degrading structural simplicity.  The value function
(Eq.~\ref{eq:value}) with $\lambda > 0$ further penalises large mutations,
ensuring that only targeted, structurally justified proposals achieve high
scores.

\paragraph{Sparse KG challenges.}
Our curated KGs range from 41--253 entities and 39--814 edges across all seven
domains (Table~\ref{tab:reproducibility}).  The smaller hand-curated domains
(41--49 entities) represent early-stage scientific KG construction where
sparsity limits the generativity component: DistMult requires $\geq 10$
training edges to produce meaningful predictions, and the 20\% masking protocol
leaves few test edges for evaluation on these domains.  The Wikidata domains
(252--253 entities, 806--814 edges) confirm that the framework scales to
medium-sized KGs.

\paragraph{Proposal quality vs.\ validity rate.}
The stagnation recovery mechanism (constrained prompting after $S = 5$
generations without valid proposals) effectively maintains a validity rate
$\geq 0.50$ across domains.  However, constrained proposals tend to cluster in
low-novelty regions of the MAP-Elites grid.  A promising direction is adaptive
constraint relaxation, where the degree of structural constraint is modulated by
archive coverage rather than a binary switch.

\paragraph{Symmetry redesign rationale.}
In the initial design, symmetry measured inter-type behavioural uniformity via
pairwise JS divergence across entity types.  Reviewer feedback correctly
identified that this penalises natural functional specialisation: in an
astronomy KG, stars and planets \emph{should} exhibit different edge-type
distributions.  The revised design (Eq.~\ref{eq:symm}) measures \emph{intra-type}
behavioural consistency---whether entities of the same type use edge types
consistently---which is a more meaningful structural signal.  Entities with no
outgoing edges are excluded, and single-entity types receive maximal consistency
by convention.

\paragraph{Contradicts validity.}
\texttt{Contradicts} edges need not represent noise---in scientific discourse,
competing hypotheses are valuable and their explicit representation is a feature,
not a defect.  Our coherence penalty targets only \emph{dense} contradiction
(high contradicts-to-edge ratio), which signals structural noise; sparse
contradiction is tolerated.  Future work includes domain-adaptive weighting,
where component weights are learned per domain via held-out validation
performance.

\paragraph{Frequency dominance and hybrid analysis.}
The frequency heuristic is highly competitive and achieves the highest Hits@10
on two of five domains (Section~\ref{sec:freq-analysis}).  This is not surprising: in small KGs with
peaked edge-type distributions, predicting the dominant type is a strong
baseline.  Rather than viewing this as a failure, we characterise the
\emph{regime} where Harmony adds complementary value
(Section~\ref{sec:regime}).  The optional $\varepsilon$ hybrid parameter
(Eq.~\ref{eq:harmony}) enables principled combination when appropriate, though
we emphasise that pure Harmony ($\varepsilon = 0$) remains the primary
contribution.

\paragraph{Harmony--downstream disconnect.}
The correlation analysis (Section~\ref{sec:correlation}) examines which Harmony
components predict downstream Hits@10 improvements.  Where components show weak
correlation, this indicates a gap between the structural quality signal and
link-prediction performance.  We view this as an important finding that
motivates future metric refinement rather than a fatal flaw: the Harmony metric
captures structural desiderata (simplicity, consistency, predictability) that
are valuable for theory assessment beyond link prediction alone.

\paragraph{Per-domain failure analysis.}
On the Physics domain, Harmony-guided proposals underperform DistMult-alone.
Failure mode analysis reveals that the LLM tends to propose edges between
highly connected hub entities (force, energy) rather than peripheral concepts,
producing proposals that are valid but redundant with existing structure.  This
suggests that domain-specific prompting strategies or proposal diversity
mechanisms may be needed for hub-dominated KGs.

\paragraph{LLM dependence and safety.}
The proposal quality depends on the LLM's domain knowledge and instruction
following.  Our experiments use a single model (\texttt{gpt-oss:20b}, a dense
20B-parameter open-source model); ensembling across model families could improve
diversity and robustness.  The island-model architecture naturally supports
heterogeneous LLM backends per island.  To mitigate the risk of LLM-generated
misinformation entering scientific workflows, proposals enter a \emph{staging
layer}: they are scored by the Harmony metric and archived, but never
automatically integrated into the base KG.  Every proposal requires an explicit
falsification condition, enabling principled rejection.  We recommend
domain-expert validation as a mandatory step in any deployment; the backtesting
results (Section~\ref{sec:results}) provide automated external validity evidence
but do not replace human judgement.

\paragraph{Scalability.}
The Harmony framework's computational cost is dominated by DistMult training
($O(|E| \cdot d \cdot \text{epochs})$) and LLM inference ($O(T_{\max} \cdot
4)$ calls for 4 islands).  The three graph-structural components
(compressibility, coherence, symmetry) are $O(|V| + |E|)$ each.  For our
hand-curated KGs (41--49 entities, 39--53 edges), total wall time is $\sim$10
minutes per domain on a single CPU.  The Wikidata domains (252--253 entities,
806--814 edges) complete in $\sim$27 minutes per seed (Table~\ref{tab:appendix_mlx_runtime}).
Scaling to medium-size KGs ($\sim$1{,}000 entities) increases DistMult training
time linearly with $|E|$ but does not change the LLM call count, making the
framework practical without GPU hardware.

\paragraph{Broader impacts.}
This work aims to accelerate scientific theory discovery by automating the
generation and evaluation of structural hypotheses in knowledge graphs.  On the
positive side, this could reduce the time researchers spend formulating initial
hypotheses and help surface non-obvious connections across disciplinary
boundaries.  On the negative side, LLM-generated proposals can be
plausible-sounding yet factually incorrect; deploying such proposals without
expert validation risks propagating erroneous claims into downstream scientific
workflows.  We mitigate this by including falsification conditions in every
proposal and recommending domain-expert validation before any claim is treated
as established.  The backtesting protocol provides automated external validity
checks but does not replace human judgement.

\paragraph{Limitations.}
(i)~The seven-relation type vocabulary, while sufficient for our seven domains,
may be too coarse for highly specialised fields (e.g.\ organic chemistry
reaction types).
(ii)~The Harmony metric treats all edge types equally in the compressibility
and coherence components; domain-specific type hierarchies could improve these
signals.
(iii)~Although we report mean $\pm$ std across 10 seeds with paired bootstrap
CIs and Cliff's delta effect sizes, the observed differences between Harmony
and DistMult-alone are small relative to cross-seed variance on smaller KGs;
larger-scale experiments with more generations or larger KGs may be needed to
establish stronger statistical significance.
(iv)~The frequency heuristic outperforms all embedding-based methods on
Hits@10 across all domains; the regime characterisation
(Section~\ref{sec:regime}) identifies conditions under which Harmony adds
complementary value, but does not demonstrate universal superiority.
(v)~Evaluation circularity between DistMult in the generativity component and
DistMult in the evaluation protocol is partially addressed by Harmony-3
($\delta = 0$) and external KGE baselines (TransE, RotatE, ComplEx), but the
primary Harmony configuration still includes DistMult in the objective.
