\section{Experiments}\label{sec:experiments}

\subsection{Knowledge Graph Domains}

We evaluate on seven KGs spanning scientific disciplines: five hand-curated from
textbook knowledge and two automatically extracted from Wikidata.  All KGs use
the shared seven-relation type vocabulary (Section~\ref{sec:schema}).

\begin{itemize}[nosep]
  \item \textbf{Linear algebra}: 17 entities (matrix, vector, eigenvalue,
    determinant, rank, etc.) with algebraic dependency and derivation edges.
  \item \textbf{Periodic table}: 22 entities (chemical elements, periods,
    groups, and categories) with trends, groups, and reactivity relations.
  \item \textbf{Astronomy}: celestial objects (star, planet, black hole, nebula)
    and astrophysical processes.
  \item \textbf{Physics}: fundamental concepts (force, energy, momentum,
    gravity) and their theoretical inter-relations.
  \item \textbf{Materials science}: material properties, compounds, and
    structure--property relationships.
  \item \textbf{Wikidata Physics}: 253 entities and 800+ edges extracted from
    Wikidata covering physical concepts, laws, and phenomena.
  \item \textbf{Wikidata Materials}: 283 entities and 900+ edges extracted from
    Wikidata covering materials, properties, and applications.
\end{itemize}

The first two domains serve as \emph{calibration} targets (known structure for
gate validation); the remaining five are \emph{discovery} targets.  The Wikidata
domains provide larger, independently sourced KGs that test scalability beyond
hand-curated data.  Table~\ref{tab:reproducibility} in the appendix provides
full dataset statistics including entity counts, edge counts, type counts, and
data sources for all seven domains.

\subsection{Dataset Splitting}

For each KG, we first reserve 10\% of edges as a hidden backtesting set,
withheld from all metric computations and proposal generation.  The remaining
90\% are split 80/10/10 into training, validation, and test sets (yielding
effective proportions of approximately 72/9/9/10 over all edges).  The
validation set is used for early stopping of DistMult training (patience of 10
epochs monitoring validation Hits@10) to prevent overfitting on small KGs.
This provides an unbiased evaluation of generativity on unseen edges.

\subsection{Baselines}

We compare Harmony-guided proposals against three heuristic baselines and three
additional KGE models.  All embedding-based methods use the same edge splits,
embedding dimension ($d = 50$), and training protocol:

\begin{enumerate}[nosep]
  \item \textbf{Random}: propose edges between random entity pairs with random
    relation types.
  \item \textbf{Frequency}: propose the most frequent relation type between the
    most-connected entity pairs.
  \item \textbf{DistMult-alone}: use DistMult's own top-ranked predictions
    without Harmony scoring or LLM involvement.
  \item \textbf{TransE} \citep{bordes2013transe}: translational distance model.
  \item \textbf{RotatE} \citep{sun2019rotate}: rotation-based model in complex
    space.
  \item \textbf{ComplEx}: complex-valued bilinear model with Hermitian product.
\end{enumerate}

\noindent To break evaluation circularity (DistMult appears in both the Harmony
objective and evaluation), we additionally report \textbf{Harmony-3}: Harmony
scoring with $\delta = 0$ (no generativity), using only compressibility,
coherence, and symmetry.  This demonstrates whether the structural components
alone provide useful signal independent of DistMult.

\subsection{Factor Decomposition}

To isolate the contributions of the LLM proposer, Harmony scoring, and the
MAP-Elites quality-diversity archive, we run three ablated configurations:

\begin{enumerate}[nosep]
  \item \textbf{LLM-only}: the LLM proposer generates proposals, but Harmony
    scoring is bypassed---all valid proposals are accepted into the archive.
  \item \textbf{Harmony-only}: a random proposer generates edges (no LLM), and
    Harmony scoring selects among them.
  \item \textbf{No-QD}: the full LLM + Harmony pipeline, but with a greedy
    single-bin archive instead of the $5 \times 5$ MAP-Elites grid.
\end{enumerate}

\subsection{Evaluation Protocol}

\paragraph{Quantitative metrics.}
We report Hits@10, Hits@3, Hits@1, and Mean Reciprocal Rank (MRR):
\begin{equation}\label{eq:mrr}
  \text{MRR} = \frac{1}{|Q|} \sum_{i=1}^{|Q|} \frac{1}{\text{rank}_i},
\end{equation}
where $Q$ is the set of masked test edges and $\text{rank}_i$ is the rank of
the true target entity among all candidates.  Metrics are computed on the test
split after applying top proposals from the MAP-Elites archive to the base KG.
All main results (Table~\ref{tab:metrics}) are averaged over 10 random seeds
($s \in \{42, 123, 456, 789, 1024, 1337, 2048, 3141, 4096, 5000\}$) for
dataset splitting, model initialisation, and edge masking.  LLM proposals are
generated by \texttt{gpt-oss:20b}, a dense 20B-parameter open-source model
locally served via Ollama with deterministic temperature settings per island.
The model is a Llama-architecture derivative fine-tuned for instruction following;
all inference runs on a single Apple M2 Pro CPU without GPU acceleration.

\paragraph{Calibration gate.}
Before running discovery experiments, we verify on the two calibration domains
(linear algebra, periodic table) that: (i)~Harmony mean $\geq 10\%$ above the
frequency baseline, and (ii)~the bootstrap 95\% CI lower bound exceeds the
frequency mean, across six pre-registered weight configurations ($\alpha \in
\{0.3, 0.5, 0.7\}$, $\beta \in \{0.1, 0.3\}$, $\gamma = \delta = 0.25$; each
vector is renormalised to sum to~1 before scoring).

\paragraph{Backtesting.}
For each domain, we evaluate top-$N$ archive proposals against the 10\%
held-out hidden edges, reporting Precision@$N$ and Recall@$N$ ($N \in
\{5, 10, 20\}$).  A proposal matches a hidden edge if its source entity, target
entity, and edge type exactly match a withheld triple.  This provides external
validity evidence independent of the DistMult training objective.

\paragraph{Archive diversity.}
We report MAP-Elites coverage (fraction of occupied cells in the $5 \times 5$
grid), best and mean fitness, and qualitative inspection of proposals across
behavioural descriptor bins.
