\section{Introduction}\label{sec:intro}

Knowledge graphs (KGs) organise scientific knowledge as typed, directed
multigraphs: entities represent concepts (e.g.\ \emph{photon},
\emph{eigenvalue}, \emph{graphene}) and edges encode semantic relations such as
\texttt{derives}, \texttt{explains}, or \texttt{contradicts}
\citep{ji2021kgsurvey}.  Despite decades of curation, scientific KGs remain
structurally incomplete---missing edges that encode latent theoretical
connections and missing entities that represent undiscovered concepts.

Knowledge graph completion (KGC) methods---TransE \citep{bordes2013transe},
DistMult \citep{yang2015distmult}, RotatE \citep{sun2019rotate}---learn
low-dimensional embeddings and predict missing links.  However, they operate at
the \emph{triple} level: each predicted link is an isolated statistical
extrapolation without theoretical justification.  They do not produce
\emph{theory-level propositions} that articulate \emph{why} a relation should
hold, what it implies, or how it could be falsified.

We address this gap with \textbf{Harmony}, a framework for automated theory
discovery in scientific KGs.  The key idea is a composite quality metric---the
\emph{Harmony score}---that captures four desiderata of a well-structured
knowledge graph:
\begin{enumerate}[nosep]
  \item \textbf{Compressibility}: the KG's edge-type distribution and spanning
    structure admit a short description (MDL proxy).
  \item \textbf{Coherence}: closed paths exhibit consistent edge-type semantics
    and contradictions are sparse.
  \item \textbf{Symmetry}: entities of the same type use edge types in similar
    proportions (low Jensen--Shannon divergence).
  \item \textbf{Generativity}: a shallow DistMult model can recover masked edges,
    indicating learnable relational patterns.
\end{enumerate}

An LLM proposes candidate mutations---adding edges or entities---each
accompanied by a claim, justification, and falsification condition.  Proposals
are scored by Harmony gain and archived in a MAP-Elites
\citep{mouret2015mapelites} quality-diversity grid across an island-model
\citep{whitley1997island} search topology (Section~\ref{sec:search}).

\paragraph{Contributions.}
\begin{enumerate}[nosep]
  \item A four-component \textbf{Harmony metric} that provides interpretable
    selection pressure (compressibility, coherence, symmetry, generativity),
    is bounded in $[0,1]$, and decomposes into independently computable
    sub-scores (Section~\ref{sec:metric}).
  \item A \textbf{proposal schema} that elevates KG mutations from bare triples
    to falsifiable theory-level claims with claim, justification, and
    falsification fields (Section~\ref{sec:proposals}).
  \item An \textbf{island-model LLM search loop} with MAP-Elites quality-diversity
    archiving and stagnation-triggered constrained prompting
    (Section~\ref{sec:search}); factor decomposition identifies the
    archive (grid + diversity bookkeeping combined) as the dominant driver of
    structural gain in our experiments.
  \item Evaluation on \textbf{seven KG domains} (five hand-curated, two
    Wikidata-sourced) with 10-seed cross-validation: Harmony shows
    directional but non-significant improvements over DistMult-alone on
    three of five discovery domains
    ($|\delta_C| \leq 0.12$, in the ``negligible'' range per
    \citet{cliff1993dominance}).  The LLM proposer's identifiable
    contribution is qualitative (semantic structure, falsifiability)
    rather than quantitative (link-prediction gain) (Section~\ref{sec:results}).
\end{enumerate}
